---
title: "ESM 206 Lab 6"
author: "Sarah Lam"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based text example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test). 

H0: Equal ranks (or medians)
HA: Ranks are not equal (or, median are not equal)

### Create samples

`set.seed()` to create "psuedorandom, but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distribution using `hist()`. 

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results: 

```{r}
ttest <- t.test(gp_1, gp_2)

#If these data were paired, I can just ass the argument 'paired = "TRUE"' within the t.test() function. 
```

This p-value means that IF our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is a `r round(ttest$p.value, 2)` probability of finding sample means *at least as different as mine* by random chance. 

How we actually relay rejecting/retaining the null hypothesis:
[A more comprehensive explanation of difference between groups ... actually means, data spread, effect size] ... there is no significant difference (t(df) = t-statistic value, p = ). 

### Mann Whitney U rank based test 

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that IF our samples are drawn from a population with the same median, there is a `r round(mwu$p.value, 2)` probability of finding sample medians *at least as different as ours* by random chance.

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly (Mann Whitney U, ...). 

### Simple linear regression

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense.

Use the `penguins` data from {palmerpenguins}, exploring flipper length (x) versus body mass (y). 

```{r}
# use View9penguins) to look at it 
# Make and exploratory scatter plot of penguin flipper length (x-axis) versus body mass (y-axis) 

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Find the model estimates (coefficnets) 

Use the `lm` function fo rlinear model. 

body_mass_g ~ <<<< tilde (nont minus sign)

```{r}
penguin_lm <- lm(data = penguins, body_mass_g ~ flipper_length_mm)
```

Model coefficients: `r penguin_lm$coefficients` 
Flipper length coefficient: `r round(penguin_lm$coefficients[2],2)`

### Use `broom::tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```

### Use broom::glance() to get output of mdoel fit & statistical significance 

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use equatiomatic::extract() to return the model in LaTex 

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use plot() function to automatically produe diagnostic plots

```{r}
plot(penguin_lm)
```


